---
title: "E.Coli"
author: "Sara Taheri"
date: "8/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
```


```{r}
data <- read.csv("Expression_data.csv")
```


```{r}
head(data)
```

Extract the related rows:

```{r}
small_data <- data %>% 
  filter(log.TPM %in% 
                c("b3067" ,"b3357","b3961","b0683", "b1334", "b1221", "b4124", "b0620", "b0564", "b0928",  "b0619", "b0972", "b0973", "b0977"))%>% 
    mutate(name = c("appY", "dpiB" , "dpiA", "fur" , "aspC"  , "hyaA" , "hyaB" , "hyaF" , "narL" , "fnr" , "rpoD"  , "crp" , "oxyR" , "dcuR"),
         label = c("b0564","b0619" ,"b0620","b0683", "b0928", "b0972", "b0973", "b0977", "b1221", "b1334", "b3067" , "b3357", "b3961", "b4124")
         ) %>%
  relocate(name) %>% 
  relocate(label)
#rownames(small_data) <- c("rpoD" , "crp" , "oxyR", "fur" , "fnr"  , "narL" , "dcuR" , "dpiA" , "appY" , "aspC"  , "dpiB" , "hyaA" , "hyaB" , "hyaF")
small_data
```



```{r}
intv_data1 <- small_data %>% select(name, fur__delfur_dpd__1, fur__delfur_dpd__2, fur__delfur_fe2__1, fur__delfur_fe2__2)
rpoD = intv_data1 %>% slice(1) %>% unlist(., use.names=FALSE)
crp = intv_data1 %>% slice(2) %>% unlist(., use.names=FALSE)
oxyR = intv_data1 %>% slice(3) %>% unlist(., use.names=FALSE)
fur = intv_data1 %>% slice(4) %>% unlist(., use.names=FALSE)
fnr = intv_data1 %>% slice(5) %>% unlist(., use.names=FALSE)
narL = intv_data1 %>% slice(6) %>% unlist(., use.names=FALSE)
dcuR = intv_data1 %>% slice(7) %>% unlist(., use.names=FALSE)
dpiA = intv_data1 %>% slice(8) %>% unlist(., use.names=FALSE)
appY = intv_data1 %>% slice(9) %>% unlist(., use.names=FALSE)
aspC = intv_data1 %>% slice(10) %>% unlist(., use.names=FALSE)
dpiB = intv_data1 %>% slice(11) %>% unlist(., use.names=FALSE)
hyaA = intv_data1 %>% slice(12) %>% unlist(., use.names=FALSE)
hyaB = intv_data1 %>% slice(13) %>% unlist(., use.names=FALSE)
hyaF = intv_data1 %>% slice(14) %>% unlist(., use.names=FALSE)
intv_data <- data.frame("rpoD" = rpoD, "crp" = crp, "oxyR" = oxyR, "fur" = fur,
                        "fnr" = fnr, "narL" = narL, "dcuR" = dcuR, "dpiA" = dpiA,
                        "appY" = appY, "aspC" = aspC, "dpiB" = dpiB, "hyaA" = hyaA,
                        "hyaB" = hyaB, "hyaF" = hyaF)
intv_data <- intv_data[-1,]
#intv_data <- intv_data[-2,]
intv_data <- apply(intv_data[ , ], 2,function(x) as.numeric(as.character(x)))
intv_data <- as.data.frame(intv_data)
intv_data
```

```{r}
obs_data1 <- small_data %>% select(-c(name,
                                      fur__delfur_dpd__1,
                                      fur__delfur_dpd__2,
                                      fur__delfur_fe2__1,
                                      fur__delfur_fe2__2,
                                      crp__delcrp_glyc__3,
                                      ica__bw_delpurR_cytd__2,
                                      pal__tartr_ale28__1, #not interventional
                                      ica__bw_delpurR_cytd__2, #not interventional
                                      ssw__glc_ac_glc1__1, #not interventional
                                      X42c__wt_42c__1,#not interventional
                                      rpoB__rpoBE672K_lb__2,#not interventional
                                      rpoB__rpoBE672K_glc__2,#not interventional
                                      rpoB__rpoBE672K_glc__1,#not interventional
                                      rpoB__rpoBE546V_lb__1#not interventional
                                      ))
rpoD = obs_data1 %>% slice(1) %>% unlist(., use.names=FALSE)
crp = obs_data1 %>% slice(2) %>% unlist(., use.names=FALSE)
oxyR = obs_data1 %>% slice(3) %>% unlist(., use.names=FALSE)
fur = obs_data1 %>% slice(4) %>% unlist(., use.names=FALSE)
fnr = obs_data1 %>% slice(5) %>% unlist(., use.names=FALSE)
narL = obs_data1 %>% slice(6) %>% unlist(., use.names=FALSE)
dcuR = obs_data1 %>% slice(7) %>% unlist(., use.names=FALSE)
dpiA = obs_data1 %>% slice(8) %>% unlist(., use.names=FALSE)
appY = obs_data1 %>% slice(9) %>% unlist(., use.names=FALSE)
aspC = obs_data1 %>% slice(10) %>% unlist(., use.names=FALSE)
dpiB = obs_data1 %>% slice(11) %>% unlist(., use.names=FALSE)
hyaA = obs_data1 %>% slice(12) %>% unlist(., use.names=FALSE)
hyaB = obs_data1 %>% slice(13) %>% unlist(., use.names=FALSE)
hyaF = obs_data1 %>% slice(14) %>% unlist(., use.names=FALSE)

obs_data <- data.frame("rpoD" = rpoD, "crp" = crp, "oxyR" = oxyR, "fur" = fur,
                       "fnr" = fnr, "narL" = narL, "dcuR" = dcuR, "dpiA" = dpiA,
                       "appY" = appY, "aspC" = aspC, "dpiB" = dpiB, "hyaA" = hyaA,
                       "hyaB" = hyaB, "hyaF" = hyaF)
obs_data <- obs_data[-1,]
obs_data <- apply(obs_data[ , ], 2,function(x) as.numeric(as.character(x)))
obs_data <- as.data.frame(obs_data)
obs_data <- obs_data[-1,]
head(obs_data)
```

Removing the outliers:

```{r}
summary(obs_data$dpiA)
obs_data <- obs_data %>% filter(dpiA < 4)
summary(obs_data$dpiA)
```

```{r}
#write.csv(obs_data, "obs_data_ecoli_without_zeros.csv")
```

# Estimating ATE from data

```{r}
mean(intv_data[c(1,3,4),"dpiA"])
```

# Estimation of ATE with linear regression BN and GAN simulated data and over real data

Here I tried to estimate ATE for a couple of examples using the real data:

```{r}
lm_dpiA_Given_fur_crp_rpoD <- lm(dpiA ~ fur + crp + rpoD, obs_data_remove_outlier)
lm_dpiA_Given_fur_crp_rpoD
b_dpiA_Given_fur_crp_rpoD <- coef(lm_dpiA_Given_fur_crp_rpoD) ['fur']

dpiA_Given_fur_crp_rpoD <- coef(lm_dpiA_Given_fur_crp_rpoD) ['(Intercept)'] + 
                           coef(lm_dpiA_Given_fur_crp_rpoD) ['crp'] * obs_data_remove_outlier$crp +
                           coef(lm_dpiA_Given_fur_crp_rpoD) ['rpoD'] * obs_data_remove_outlier$rpoD
summary(dpiA_Given_fur_crp_rpoD)
var(dpiA_Given_fur_crp_rpoD)
```

```{r}
lm_dcuR_Given_fur_crp_rpoD_narL <- lm(dcuR ~ fur + crp + rpoD + narL, obs_data_remove_outlier)
lm_dcuR_Given_fur_crp_rpoD_narL
dpiA_Given_fur_crp_rpoD_narL <- coef(lm_dcuR_Given_fur_crp_rpoD_narL) ['(Intercept)'] + 
                                coef(lm_dcuR_Given_fur_crp_rpoD_narL) ['crp'] * obs_data_remove_outlier$crp +
                                coef(lm_dcuR_Given_fur_crp_rpoD_narL) ['rpoD'] * obs_data_remove_outlier$rpoD +
                                coef(lm_dcuR_Given_fur_crp_rpoD_narL) ['narL'] * obs_data_remove_outlier$narL
summary(dpiA_Given_fur_crp_rpoD_narL)
var(dpiA_Given_fur_crp_rpoD_narL)
```


```{r}
lm_dcuR_Given_fur_crp_rpoD_oxyR <- lm(dcuR ~ fur + crp + rpoD + oxyR, obs_data)
dcuR_Given_fur_crp_rpoD_oxyR <- coef(lm_dcuR_Given_fur_crp_rpoD_oxyR) ['(Intercept)'] + 
                                coef(lm_dcuR_Given_fur_crp_rpoD_oxyR) ['crp'] * obs_data_remove_outlier$crp +
                                coef(lm_dcuR_Given_fur_crp_rpoD_oxyR) ['rpoD'] * obs_data_remove_outlier$rpoD +
                                coef(lm_dcuR_Given_fur_crp_rpoD_oxyR) ['oxyR'] * obs_data_remove_outlier$narL
summary(dcuR_Given_fur_crp_rpoD_oxyR)
var(dcuR_Given_fur_crp_rpoD_oxyR)
```

Here I got the results for running the linear case for the GAN and BN approach using Algorithm 1 in the paper:

```{r}
BN_pathwise <- readRDS("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/BN_SimulatedResults/BN_Path_furfnrdpiA_results.rdata")
summary_path <- BN_pathwise %>% group_by(formula,stage_1_formula) %>% summarise(N=n(),mean=mean(Ey_dox),sd=sd(Ey_dox)) %>% arrange(sd)
BN_backdoor <- read.csv("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/BN_SimulatedResults/BN_SimulatedDatasets_Backdoor_EdpiA_dofur0.csv")

GAN_pathwise <- readRDS("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/GAN/GAN_Path_furfnrdpiA_results.rdata")
GAN_backdoor <- read.csv("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/GAN/gan_backdoor.csv")
```

```{r}
# (a) full network
lm_BN_result_full <- BN_pathwise %>% filter(formula == "dpiA ~ fnr + crp + rpoD + oxyR + dcuR + narL + fur" & stage_1_formula == "fnr ~ fur + rpoD")
lm_GAN_result_full <- GAN_pathwise %>% filter(formula == "dpiA ~ fnr + crp + rpoD + oxyR + dcuR + narL + fur" & stage_1_formula == "fnr ~ fur + rpoD")
# (b) olive
lm_BN_result_olive <- BN_backdoor %>% filter(formula == "dpiA ~ fur + crp + rpoD")
lm_GAN_result_olive <- GAN_backdoor %>% filter(formula == "dpiA ~ fur + crp + rpoD")
# (c) Olive + fnr
lm_BN_result_olive_fnr <- BN_pathwise %>% filter(formula == "dpiA ~ fnr + fur + rpoD + dcuR + crp" & stage_1_formula == "fnr ~ fur + rpoD + crp")
lm_GAN_result_olive_fnr <- GAN_pathwise %>% filter(formula == "dpiA ~ fnr + fur + rpoD + dcuR + crp" & stage_1_formula == "fnr ~ fur + rpoD + crp")
# (d) Olive + green
lm_BN_result_olive_green <- BN_pathwise %>% filter(formula == "dpiA ~ fnr + fur + rpoD + dcuR + narL + crp" & stage_1_formula == "fnr ~ fur + rpoD")
lm_GAN_result_olive_green <- GAN_pathwise %>% filter(formula == "dpiA ~ fnr + fur + rpoD + dcuR + narL + crp" & stage_1_formula == "fnr ~ fur + rpoD")
```


# Causal generative approach

## Full model and also the case with all the mediators (Fig 8 (a) and (d))

```{r}
model_str_EColi <- "
    data {
        int D;
        vector[D] rpoD;
        vector[D] crp;
        vector[D] fur;
        vector[D] fnr;
        vector[D] narL;
        vector[D] dcuR;
        vector[D] dpiA;
    }
  parameters {
       real<lower=0> mu_rpoD; 
       real<lower=0> sigma_rpoD;
       
       real<lower=-2, upper=2> beta0_crp;
       real<lower=-2, upper=2> beta_rpoDTocrp;
       
       real<lower=-2, upper=2> beta0_fnr;
       real<lower=-2, upper=2> beta_furTofnr;
       real<lower=-2, upper=2> beta_rpoDTofnr;
       
       real<lower=-2, upper=2> beta0_narL;
       real<lower=-2, upper=2> beta_fnrTonarL;
       real<lower=-2, upper=2> beta_rpoDTonarL;
       
       real<lower=-2, upper=2> beta0_dcuR;
       real<lower=-2, upper=2> beta_crpTodcuR;
       real<lower=-2, upper=2> beta_fnrTodcuR;
       real<lower=-2, upper=2> beta_rpoDTodcuR;
       real<lower=-2, upper=2> beta_narLTodcuR;
       
       real<lower=-2, upper=2> beta0_dpiA;
       real<lower=-2, upper=2> beta_crpTodpiA;
       real<lower=-2, upper=2> beta_rpoDTodpiA;
       real<lower=-2, upper=2> beta_fnrTodpiA;
       real<lower=-2, upper=2> beta_dcuRTodpiA;
       real<lower=-2, upper=2> beta_narLTodpiA;
    }
    transformed parameters {
      vector[D] crp_loc;
      vector[D] fnr_loc;
      vector[D] narL_loc;
      vector[D] dcuR_loc;
      vector[D] dpiA_loc;
      for (i in 1:D){
        crp_loc[i] = 6 / (1 + exp(-beta0_crp - rpoD[i] * beta_rpoDTocrp));
        fnr_loc[i] = 11.57 / (1 + exp(-beta0_fnr - fur[i] * beta_furTofnr - rpoD[i] * beta_rpoDTofnr));
        narL_loc[i] = 3.7 / (1 + exp(-beta0_narL - fnr[i] * beta_fnrTonarL - rpoD[i] * beta_rpoDTonarL));
        dcuR_loc[i] = 4.2 / (1 + exp(-beta0_dcuR - crp[i] * beta_crpTodcuR - fnr[i] * beta_fnrTodcuR - rpoD[i] * beta_rpoDTodcuR - narL[i] * beta_narLTodcuR));
        dpiA_loc[i] = 3.8 / (1 + exp(-beta0_dpiA - crp[i] * beta_crpTodpiA - fnr[i] * beta_fnrTodpiA - rpoD[i] * beta_rpoDTodpiA - narL[i] * beta_narLTodpiA - dcuR[i] * beta_dcuRTodpiA));
      }
    }
    model {
        mu_rpoD ~ normal(5.7, 1);
        sigma_rpoD ~ normal(1.6,1);
        
        beta0_crp ~ normal(0,10);
        beta_rpoDTocrp ~ normal(0,10);
        
        beta0_fnr ~ normal(0,10);
        beta_furTofnr ~ normal(0,10);
        beta_rpoDTofnr ~ normal(0,10);
        
        beta0_narL ~ normal(0,10);
        beta_fnrTonarL ~ normal(0,10);
        beta_rpoDTonarL ~ normal(0,10);
        
        beta0_dcuR ~ normal(0,10);
        beta_crpTodcuR ~ normal(0,10);
        beta_fnrTodcuR ~ normal(0,10);
        beta_rpoDTodcuR ~ normal(0,10);
        beta_narLTodcuR ~ normal(0,10);
        
        beta0_dpiA ~ normal(0,10);
        beta_crpTodpiA ~ normal(0,10);
        beta_rpoDTodpiA ~ normal(0,10);
        beta_fnrTodpiA ~ normal(0,10);
        beta_dcuRTodpiA ~ normal(0,10);
        beta_narLTodpiA ~ normal(0,10);
        
        rpoD ~ normal(mu_rpoD, sigma_rpoD);
        crp ~ normal(crp_loc,5);
        fnr ~ normal(fnr_loc,0.5);
        narL ~ normal(narL_loc,0.7);
        dcuR ~ normal(dcuR_loc,0.76);
        dpiA ~ normal(dpiA_loc,0.6);
    }
"
```

```{r, message=FALSE, warning=FALSE}
mod_EColi <- rstan::stan_model(model_code = model_str_EColi)
```

```{r}
BN_datasets <- readRDS("Experimental_case_study/BN_SimulatedDatasets.RData")
obs_data_list <- BN_datasets
obs_data_list[[length(BN_datasets)+1]] <- obs_data
```

```{r}
GAN_datasets <- list()
for (i in 1:100) {
  GAN_datasets[[i]] <- read.csv(paste("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/GAN/GAN_simulated data/simdata",i,".csv",sep=""))
}
obs_data_list <- GAN_datasets
obs_data_list[[length(BN_datasets)+1]] <- obs_data
```


```{r, echo = FALSE}
K=100
ndpu = nrow(obs_data)
hmc_fit_list <- rep(list(list()), K)

start_time <- Sys.time()
for (num_data_points in ndpu) {
  for (data_count in 1:K) {
    print(data_count)
    data_list <- list(D=num_data_points,
                      rpoD = obs_data_list[[data_count]]$rpoD[1:num_data_points],
                      crp = obs_data_list[[data_count]]$crp[1:num_data_points],
                      fur = obs_data_list[[data_count]]$fur[1:num_data_points],
                      fnr = obs_data_list[[data_count]]$fnr[1:num_data_points],
                      narL = obs_data_list[[data_count]]$narL[1:num_data_points],
                      dcuR = obs_data_list[[data_count]]$dcuR[1:num_data_points],
                      dpiA = obs_data_list[[data_count]]$dpiA[1:num_data_points]
    )
    
    hmc_fit_list[[data_count]][[paste0("num_data_points_",num_data_points)]] <- rstan::sampling(mod_EColi,
                                                                                                data=data_list,
                                                                                                chains = 2,
                                                                                                iter = 3000,
                                                                                                warmup = 1500,
                                                                                                seed = 1)
  }
  
}
end_time <- Sys.time()
Time = end_time - start_time
print(Time)
#Time difference of 17.49014 mins mins total
#saveRDS(hmc_fit_list,file = "data/IGF_hmc_fit/hmc_fit_list_IGF_olive.RData")
```

### Mutilated model

```{r}
mutilated_model_EColi <- function(mu_rpoD, sigma_rpoD,
                                  beta0_crp, beta_rpoDTocrp, 
                                  beta0_fnr, beta_furTofnr, beta_rpoDTofnr,
                                  beta0_narL, beta_fnrTonarL, beta_rpoDTonarL,
                                  beta0_dcuR, beta_crpTodcuR, beta_fnrTodcuR, beta_rpoDTodcuR, beta_narLTodcuR,
                                  beta0_dpiA, beta_crpTodpiA, beta_rpoDTodpiA, beta_fnrTodpiA, beta_dcuRTodpiA, beta_narLTodpiA,
                                  fur, num_gen_samples,seed) {
  set.seed(seed)
  rpoD = rnorm(num_gen_samples, mu_rpoD, sigma_rpoD) 
  crp = 6 / (1 + exp(-beta0_crp - rpoD * beta_rpoDTocrp))
  fur = rep(fur, num_gen_samples)
  fnr = 11.57 / (1 + exp(-beta0_fnr - fur * beta_furTofnr - rpoD * beta_rpoDTofnr))
  narL = 3.7 / (1 + exp(-beta0_narL - fnr * beta_fnrTonarL - rpoD * beta_rpoDTonarL))
  dcuR = 4.2 / (1 + exp(-beta0_dcuR - crp * beta_crpTodcuR - fnr * beta_fnrTodcuR - rpoD * beta_rpoDTodcuR - narL * beta_narLTodcuR))
  dpiA = 3.8 / (1 + exp(-beta0_dpiA - crp * beta_crpTodpiA - fnr * beta_fnrTodpiA - rpoD * beta_rpoDTodpiA - narL * beta_narLTodpiA - dcuR * beta_dcuRTodpiA))
  return(mean(dpiA))
}
```

```{r}
parameters <- c("mu_rpoD", "sigma_rpoD",
                "beta0_crp", "beta_rpoDTocrp",
                "beta0_fnr", "beta_furTofnr", "beta_rpoDTofnr",
                "beta0_narL", "beta_fnrTonarL", "beta_rpoDTonarL",
                "beta0_dcuR", "beta_crpTodcuR", "beta_fnrTodcuR", "beta_rpoDTodcuR", "beta_narLTodcuR",
                "beta0_dpiA", "beta_crpTodpiA", "beta_rpoDTodpiA", "beta_fnrTodpiA", "beta_dcuRTodpiA", "beta_narLTodpiA"
                )
```

```{r, echo = FALSE}
#hmc_fit_list <- readRDS("data/IGF_hmc_fit/hmc_fit_list_IGF_olive_ndpu20.RData")

CE_list <- list()
means_hmc <- c()
J <- c()

for (num_data_points in ndpu) {
  for (data_count in 1:K) {
    samples_hmc <- rstan::extract(hmc_fit_list[[data_count]][[paste0("num_data_points_",num_data_points)]], parameters)
    for (j in 1:3000) {
      J <-  c(J,mutilated_model_EColi(mu_rpoD = (samples_hmc$mu_rpoD)[j],
                                      sigma_rpoD = (samples_hmc$sigma_rpoD)[j],
                                      beta0_crp = (samples_hmc$beta0_crp)[j],
                                      beta_rpoDTocrp = (samples_hmc$beta_rpoDTocrp)[j],
                                      
                                      beta0_fnr = (samples_hmc$beta0_fnr)[j],
                                      beta_furTofnr = (samples_hmc$beta_furTofnr)[j],
                                      beta_rpoDTofnr = (samples_hmc$beta_rpoDTofnr)[j],
                                      
                                      beta0_narL = (samples_hmc$beta0_narL)[j],
                                      beta_fnrTonarL = (samples_hmc$beta_fnrTonarL)[j],
                                      beta_rpoDTonarL = (samples_hmc$beta_rpoDTonarL)[j],
                                      
                                      beta0_dcuR = (samples_hmc$beta0_dcuR)[j],
                                      beta_crpTodcuR = (samples_hmc$beta_crpTodcuR)[j],
                                      beta_fnrTodcuR = (samples_hmc$beta_fnrTodcuR)[j],
                                      beta_rpoDTodcuR = (samples_hmc$beta_rpoDTodcuR)[j],
                                      beta_narLTodcuR = (samples_hmc$beta_narLTodcuR)[j],
                                      
                                      beta0_dpiA = (samples_hmc$beta0_dpiA)[j],
                                      beta_crpTodpiA = (samples_hmc$beta_crpTodpiA)[j],
                                      beta_rpoDTodpiA = (samples_hmc$beta_rpoDTodpiA)[j],
                                      beta_fnrTodpiA = (samples_hmc$beta_fnrTodpiA)[j],
                                      beta_dcuRTodpiA = (samples_hmc$beta_dcuRTodpiA)[j],
                                      beta_narLTodpiA = (samples_hmc$beta_narLTodpiA)[j],
                                      fur = 0,
                                      num_gen_samples = 1000,
                                      seed = 1))
      
    }
    means_hmc <- c(means_hmc,mean(J))
    J <- c()
  }
  #CE_list[[paste0("num_data_points_",num_data_points)]] <- means_hmc
  #means_hmc <- c()
}

#saveRDS(means_hmc,file = "data/IGF_hmc_fit/means_hmc_IGF_olive.RData")
#saveRDS(means_hmc,file = "means_hmc_IGF_Raf_Mek_latent.RData")
```

```{r}
summary(means_hmc)
```


## Olive nodes with only one mediator, dcuR (Fig 8 (c))

```{r}
model_str_EColi <- "
    data {
        int D;
        vector[D] rpoD;
        vector[D] crp;
        vector[D] fur;
        vector[D] fnr;
        vector[D] dpiA;
    }
  parameters {
       real<lower=0> mu_rpoD; 
       real<lower=0> sigma_rpoD;
       
       real<lower=-2, upper=2> beta0_crp;
       real<lower=-2, upper=2> beta_rpoDTocrp;
       
       real<lower=-2, upper=2> beta0_fnr;
       real<lower=-2, upper=2> beta_furTofnr;
       real<lower=-2, upper=2> beta_rpoDTofnr;
       
       real<lower=-2, upper=2> beta0_dpiA;
       real<lower=-2, upper=2> beta_crpTodpiA;
       real<lower=-2, upper=2> beta_rpoDTodpiA;
       real<lower=-2, upper=2> beta_fnrTodpiA;
       
       vector[D] U;
    }
    transformed parameters {
      vector[D] crp_loc;
      vector[D] fnr_loc;
      vector[D] dpiA_loc;
      for (i in 1:D){
        crp_loc[i]  = 6   / (1 + exp(-beta0_crp - rpoD[i] * beta_rpoDTocrp));
        fnr_loc[i] = 11 / (1 + exp(-beta0_fnr - fur[i] * beta_furTofnr - rpoD[i] * beta_rpoDTofnr));
        dpiA_loc[i] = 3.8 / (1 + exp(-beta0_dpiA - crp[i] * beta_crpTodpiA - rpoD[i] * beta_rpoDTodpiA - fnr[i] * beta_fnrTodpiA ));
      }
    }
    model {
        mu_rpoD ~ normal(5.7, 1);
        sigma_rpoD ~ normal(1.6,1);
        
        beta0_crp ~ normal(0,10);
        beta_rpoDTocrp ~ normal(0,10);
        
        beta0_fnr ~ normal(0,10);
        beta_furTofnr ~ normal(0,10);
        beta_rpoDTofnr ~ normal(0,10);
        
        beta0_dpiA ~ normal(0,10);
        beta_crpTodpiA ~ normal(0,10);
        beta_rpoDTodpiA ~ normal(0,10);
        beta_fnrTodpiA ~ normal(0,10);
        
        rpoD ~ normal(mu_rpoD, sigma_rpoD);
        crp ~ normal(crp_loc,5);
        fnr ~ normal(fnr_loc,0.57);
        dpiA ~ normal(dpiA_loc,0.6);
    }
"
```

```{r, message=FALSE, warning=FALSE}
mod_EColi <- rstan::stan_model(model_code = model_str_EColi)
```

```{r}
BN_datasets <- readRDS("BN_SimulatedDatasets.RData")
obs_data_list <- BN_datasets
obs_data_list[[length(BN_datasets)+1]] <- obs_data
```

```{r}
GAN_datasets <- list()
for (i in 1:100) {
  GAN_datasets[[i]] <- read.csv(paste("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/GAN/GAN_simulated data/simdata",i,".csv",sep=""))
}
obs_data_list <- GAN_datasets
obs_data_list[[length(BN_datasets)+1]] <- obs_data
```

```{r, echo = FALSE}
K=100
ndpu = nrow(obs_data)
hmc_fit_list <- rep(list(list()), K)

start_time <- Sys.time()
for (num_data_points in ndpu) {
  for (data_count in 1:K) {
    print(data_count)
    data_list <- list(D=num_data_points,
                      rpoD = obs_data_list[[data_count]]$rpoD[1:num_data_points],
                      crp = obs_data_list[[data_count]]$crp[1:num_data_points],
                      fur = obs_data_list[[data_count]]$fur[1:num_data_points],
                      fnr = obs_data_list[[data_count]]$fnr[1:num_data_points],
                      dpiA = obs_data_list[[data_count]]$dpiA[1:num_data_points]
    )
    
    hmc_fit_list[[data_count]][[paste0("num_data_points_",num_data_points)]] <- rstan::sampling(mod_EColi,
                                                                                                data=data_list,
                                                                                                chains = 2,
                                                                                                iter = 3000,
                                                                                                warmup = 1500,
                                                                                                seed = 1)
  }
  
}
end_time <- Sys.time()
Time = end_time - start_time
print(Time)
#Time difference of 22.0289 hours total
saveRDS(hmc_fit_list,file = "hmc_fit_list_EColi_Olive_fnr_ndpu262.RData")
```

### Mutilated model

```{r}
mutilated_model_EColi <- function(mu_rpoD, sigma_rpoD,
                                  beta0_crp, beta_rpoDTocrp, 
                                  beta0_fnr, beta_furTofnr, beta_rpoDTofnr,
                                  beta0_dpiA, beta_crpTodpiA, beta_rpoDTodpiA, beta_fnrTodpiA,
                                  fur, num_gen_samples,seed) {
  set.seed(seed)
  rpoD = rnorm(num_gen_samples, mu_rpoD, sigma_rpoD) 
  crp = 6 / (1 + exp(-beta0_crp - rpoD * beta_rpoDTocrp))
  fur = rep(fur, num_gen_samples)
  fnr = 11 / (1 + exp(-beta0_fnr - fur * beta_furTofnr - rpoD * beta_rpoDTofnr))
  dpiA = 3.8 / (1 + exp(-beta0_dpiA - crp * beta_crpTodpiA - rpoD * beta_rpoDTodpiA - fnr * beta_fnrTodpiA))
  return(mean(dpiA))
}
```

```{r}
parameters <- c("mu_rpoD", "sigma_rpoD",
                "beta0_crp", "beta_rpoDTocrp",
                "beta0_fnr", "beta_furTofnr", "beta_rpoDTofnr",
                "beta0_dpiA", "beta_crpTodpiA", "beta_rpoDTodpiA", "beta_fnrTodpiA"
                )
```

```{r, echo = FALSE}
#hmc_fit_list <- readRDS("data/IGF_hmc_fit/hmc_fit_list_IGF_olive_ndpu20.RData")

CE_list <- list()
means_hmc <- c()
J <- c()

for (num_data_points in ndpu) {
  for (data_count in 1:K) {
    samples_hmc <- rstan::extract(hmc_fit_list[[data_count]][[paste0("num_data_points_",num_data_points)]], parameters)
    for (j in 1:3000) {
      J <-  c(J,mutilated_model_EColi(mu_rpoD = (samples_hmc$mu_rpoD)[j],
                                      sigma_rpoD = (samples_hmc$sigma_rpoD)[j],
                                      
                                      beta0_crp = (samples_hmc$beta0_crp)[j],
                                      beta_rpoDTocrp = (samples_hmc$beta_rpoDTocrp)[j],
                                      
                                      beta0_fnr = (samples_hmc$beta0_fnr)[j],
                                      beta_furTofnr = (samples_hmc$beta_furTofnr)[j],
                                      beta_rpoDTofnr = (samples_hmc$beta_rpoDTofnr)[j],
                                      
                                      beta0_dpiA = (samples_hmc$beta0_dpiA)[j],
                                      beta_crpTodpiA = (samples_hmc$beta_crpTodpiA)[j],
                                      beta_rpoDTodpiA = (samples_hmc$beta_rpoDTodpiA)[j],
                                      beta_fnrTodpiA = (samples_hmc$beta_fnrTodpiA)[j],
                                      fur = 0,
                                      num_gen_samples = 1000,
                                      seed = 1))
      
    }
    means_hmc <- c(means_hmc,mean(J))
    J <- c()
  }
  #CE_list[[paste0("num_data_points_",num_data_points)]] <- means_hmc
  #means_hmc <- c()
}

#saveRDS(means_hmc,file = "data/IGF_hmc_fit/means_hmc_IGF_olive.RData")
#saveRDS(means_hmc,file = "means_hmc_IGF_Raf_Mek_latent.RData")
```
## Olive nodes only, (Fig 8 (b))

```{r}
model_str_EColi <- "
    data {
        int D;
        vector[D] rpoD;
        vector[D] crp;
        vector[D] fur;
        vector[D] dpiA;
    }
  parameters {
       real<lower=0> mu_rpoD; 
       real<lower=0> sigma_rpoD;
       
       real<lower=-2, upper=2> beta0_crp;
       real<lower=-2, upper=2> beta_rpoDTocrp;
       
       
       real<lower=-2, upper=2> beta0_dpiA;
       real<lower=-2, upper=2> beta_crpTodpiA;
       real<lower=-2, upper=2> beta_rpoDTodpiA;
       real<lower=-2, upper=2> beta_furTodpiA;
       
       vector[D] U;
    }
    transformed parameters {
      vector[D] crp_loc;
      vector[D] dpiA_loc;
      for (i in 1:D){
        crp_loc[i]  = 2.6   / (1 + exp(-beta0_crp - rpoD[i] * beta_rpoDTocrp));
        dpiA_loc[i] = 3.8 / (1 + exp(-beta0_dpiA - crp[i] * beta_crpTodpiA - rpoD[i] * beta_rpoDTodpiA - fur[i] * beta_furTodpiA ));
      }
    }
    model {
        mu_rpoD ~ normal(5.7, 1);
        sigma_rpoD ~ normal(1.6,1);
        
        beta0_crp ~ normal(0,10);
        beta_rpoDTocrp ~ normal(0,10);
        
        beta0_dpiA ~ normal(0,10);
        beta_crpTodpiA ~ normal(0,10);
        beta_rpoDTodpiA ~ normal(0,10);
        beta_furTodpiA ~ normal(0,10);
        
        rpoD ~ normal(mu_rpoD, sigma_rpoD);
        crp ~ normal(crp_loc,5);
        dpiA ~ normal(dpiA_loc,0.6);
    }
"
```

```{r, message=FALSE, warning=FALSE}
mod_EColi <- rstan::stan_model(model_code = model_str_EColi)
```

```{r}
# BN_datasets <- readRDS("/home/rahiminasabe/BN_SimulatedDatasets.RData")
# obs_data_list <- BN_datasets
# obs_data_list[[length(BN_datasets)+1]] <- obs_data
```

```{r}
# GAN_datasets <- list()
# for (i in 1:100) {
#   GAN_datasets[[i]] <- read.csv(paste("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/GAN/GAN_simulated data/simdata",i,".csv",sep=""))
# }
# obs_data_list <- GAN_datasets
# obs_data_list[[length(BN_datasets)+1]] <- obs_data
```

```{r}
obs_data_list <- readRDS("/Users/sarataheri/GitHub/Simplified_LVM/Experimental_case_study/GAN/GAN_simulated data/GAN_datasets.RData")
```


```{r, echo = FALSE}
K=1
ndpu = nrow(obs_data)
hmc_fit_list <- rep(list(list()), K)

start_time <- Sys.time()
for (num_data_points in ndpu) {
  for (data_count in 1:K) {
    print(data_count)
    data_list <- list(D=num_data_points,
                      rpoD = obs_data_list[[data_count]]$rpoD[1:num_data_points],
                      crp = obs_data_list[[data_count]]$crp[1:num_data_points],
                      fur = obs_data_list[[data_count]]$fur[1:num_data_points],
                      fnr = obs_data_list[[data_count]]$fnr[1:num_data_points],
                      dpiA = obs_data_list[[data_count]]$dpiA[1:num_data_points]
    )
    
    hmc_fit_list[[data_count]][[paste0("num_data_points_",num_data_points)]] <- rstan::sampling(mod_EColi,
                                                                                                data=data_list,
                                                                                                chains = 2,
                                                                                                iter = 3000,
                                                                                                warmup = 1500,
                                                                                                seed = 1)
  }
  
}
end_time <- Sys.time()
Time = end_time - start_time
print(Time)
#Time difference of 22.0289 hours total
#saveRDS(hmc_fit_list,file = "hmc_fit_list_EColi_Olive_fnr_ndpu262.RData")
```

### Mutilated model

```{r}
mutilated_model_EColi <- function(mu_rpoD, sigma_rpoD,
                                  beta0_crp, beta_rpoDTocrp, 
                                  beta0_dpiA, beta_crpTodpiA, beta_rpoDTodpiA, beta_furTodpiA,
                                  fur, num_gen_samples,seed) {
  set.seed(seed)
  rpoD = rnorm(num_gen_samples, mu_rpoD, sigma_rpoD) 
  crp = 2.6 / (1 + exp(-beta0_crp - rpoD * beta_rpoDTocrp))
  fur = rep(fur, num_gen_samples)
  dpiA = 3.8 / (1 + exp(-beta0_dpiA - crp * beta_crpTodpiA - rpoD * beta_rpoDTodpiA - fur * beta_furTodpiA))
  return(mean(dpiA))
}
```

```{r}
parameters <- c("mu_rpoD", "sigma_rpoD",
                "beta0_crp", "beta_rpoDTocrp",
                "beta0_dpiA", "beta_crpTodpiA", "beta_rpoDTodpiA", "beta_furTodpiA"
                )
```

```{r, echo = FALSE}
#hmc_fit_list <- readRDS("data/IGF_hmc_fit/hmc_fit_list_IGF_olive_ndpu20.RData")

CE_list <- list()
means_hmc <- c()
J <- c()

for (num_data_points in ndpu) {
  for (data_count in 1:K) {
    samples_hmc <- rstan::extract(hmc_fit_list[[data_count]][[paste0("num_data_points_",num_data_points)]], parameters)
    for (j in 1:3000) {
      J <-  c(J,mutilated_model_EColi(mu_rpoD = (samples_hmc$mu_rpoD)[j],
                                      sigma_rpoD = (samples_hmc$sigma_rpoD)[j],
                                      
                                      beta0_crp = (samples_hmc$beta0_crp)[j],
                                      beta_rpoDTocrp = (samples_hmc$beta_rpoDTocrp)[j],
                                      
                                      beta0_dpiA = (samples_hmc$beta0_dpiA)[j],
                                      beta_crpTodpiA = (samples_hmc$beta_crpTodpiA)[j],
                                      beta_rpoDTodpiA = (samples_hmc$beta_rpoDTodpiA)[j],
                                      beta_furTodpiA = (samples_hmc$beta_furTodpiA)[j],
                                      fur = 0,
                                      num_gen_samples = 1000,
                                      seed = 1))
      
    }
    means_hmc <- c(means_hmc,mean(J))
    J <- c()
  }
  #CE_list[[paste0("num_data_points_",num_data_points)]] <- means_hmc
  #means_hmc <- c()
}

#saveRDS(means_hmc,file = "data/IGF_hmc_fit/means_hmc_IGF_olive.RData")
#saveRDS(means_hmc,file = "means_hmc_IGF_Raf_Mek_latent.RData")
```

# Plot the results

```{r}
K=100
#results for causal generative approach with BN data
means_full_cg_BN <-  readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_full_cg_BN_ndpu262.RData")
means_olive_cg_BN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_olive_cg_BN_ndpu262.RData")
means_olive_fnr_cg_BN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_olive_fnr_cg_BN_ndpu262.RData")
means_olive_green_cg_BN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_olive_green_cg_BN_ndpu262.RData")

#results for causal generative approach with GAN data
means_full_cg_GAN <-  readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_full_cg_GAN_ndpu262.RData")
means_olive_cg_GAN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_olive_cg_GAN_ndpu262.RData")
means_olive_fnr_cg_GAN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_olive_fnr_cg_GAN_ndpu262.RData")
means_olive_green_cg_GAN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/CausalGenerative_approach/data/EColi_olive_green_cg_GAN_ndpu262.RData")

#results for linear regression approach with BN data
means_full_BN <-  readRDS("/Users/sarataheri/GitHub/Experimental_case_study/BN_SimulatedResults/lm_BN_result_full.RData")
means_olive_BN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/BN_SimulatedResults/lm_BN_result_olive.RData")
means_olive_fnr_BN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/BN_SimulatedResults/lm_BN_result_olive_fnr.RData")
means_olive_green_BN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/BN_SimulatedResults/lm_BN_result_olive_green.RData")

#results for linear regression approach with GAN data
means_full_GAN <-  readRDS("/Users/sarataheri/GitHub/Experimental_case_study/GAN/lm_GAN_result_full.RData")
means_olive_GAN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/GAN/lm_GAN_result_olive.RData")
means_olive_fnr_GAN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/GAN/lm_GAN_result_olive_fnr.RData")
means_olive_green_GAN <- readRDS("/Users/sarataheri/GitHub/Experimental_case_study/GAN/lm_GAN_result_olive_green.RData")

df       <-  data.frame("CE" = c(means_full_cg_BN,
                                 means_olive_cg_BN,
                                 means_olive_fnr_cg_BN,
                                 means_olive_green_cg_BN,
                                 means_full_cg_GAN,
                                 means_olive_cg_GAN,
                                 means_olive_fnr_cg_GAN,
                                 means_olive_green_cg_GAN,
                                 means_full_BN,
                                 means_olive_BN,
                                 means_olive_fnr_BN,
                                 means_olive_green_BN,
                                 means_full_GAN,
                                 means_olive_GAN,
                                 means_olive_fnr_GAN,
                                 means_olive_green_GAN
                                 ),
                        "N" = c(rep("(a)",K),
                                rep("(b)",K),
                                rep("(c)",K),
                                rep("(d)",K),
                                rep("(a)",K),
                                rep("(b)",K),
                                rep("(c)",K),
                                rep("(d)",K),
                                rep("(a)",K),
                                rep("(b)",K),
                                rep("(c)",K),
                                rep("(d)",K),
                                rep("(a)",K),
                                rep("(b)",K),
                                rep("(c)",K),
                                rep("(d)",K)),
                        "model" = c(rep("Causal generative (BN data)", 4*K),
                                    rep("Causal generative (GAN data)", 4*K),
                                    rep("Linear regression (BN data)", 4*K),
                                    rep("Linear regression (GAN data)", 4*K)
                                    )
                        )
```


```{r}
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}

gg1 <- df_bp_CE1 %>% ggplot(aes(x = as.factor(N), y = CE, fill=model)) +
  geom_boxplot(width = 0.6) +
  geom_abline(intercept=88.5, slope=0) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.title = element_blank(),
        legend.text = element_text(size = 35,face="bold"),
        legend.key.size = unit(2, "cm"),
        axis.text.y=element_text(size=35),
        axis.title.y=element_text(size=35,face="bold"),
        axis.title.x = element_text(size=35,face="bold"),
        axis.text.x = element_text(#angle = 90, vjust = 0.5, hjust=1,
          size = 35, face="bold"),
        legend.position = "top",
        plot.title = element_text(size = 35, hjust = 0.5, face = "bold")
  ) +
  xlab("Input network") +
  #ylim(86,89) +
  #ggtitle("IGF Signalling model") +
  ylab(expression(bold(paste(hat(E),"[ Erk | do(Akt = 80) ]"))))
gg1
#ggsave("/Users/sarataheri/GitHub/Simplified_LVM/img/IGF_boxplots.pdf", plot = gg1, width = 12, height = 8, dpi = 300, units = "in")
```